### 2018-04-29 conclusions.tex
AFL capitalization change

As future work, we would like to make the automatic detection of
values that should not be visible (e.g., count towards nondeterminism)
possible: if a value (e.g., a timestamp) is always different in every
run, it is likely of no interest to developers.  This would make
testing libraries mixing deterministic behavior and nondeterministic
behavior, e.g. cryptographic libraries offering high-quality random
number generators, easier for users, who would only be shown
``surprising'' nondeterminism.  Similarly, in order to extend the
utility of vertical nondeterminism detection, we would like to
automatically identify usually-idempotent operations, and check them
for deviation from the expected behavior.  We would also like to lower the cost
of checking process nondeterminism, perhaps by using methods taken in
afl \cite{aflfuzz} to avoid high startup costs for test executions.

>>>>

As future work, we would like to make the automatic detection of
values that should not be visible (e.g., count towards nondeterminism)
possible: if a value (e.g., a timestamp) is always different in every
run, it is likely of no interest to developers.  This would make
testing libraries mixing deterministic behavior and nondeterministic
behavior, e.g. cryptographic libraries offering high-quality random
number generators, easier for users, who would only be shown
``surprising'' nondeterminism.  Similarly, in order to extend the
utility of vertical nondeterminism detection, we would like to
automatically identify usually-idempotent operations, and check them
for deviation from the expected behavior.  We would also like to lower the cost
of checking process nondeterminism, perhaps by using methods taken in
AFL \cite{aflfuzz} to avoid high startup costs for test executions.

### 2018-04-29 experiments.tex
AFL capitalization change

Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms.  The {\tt flatten\_and\_sort} function is needed because we care about actual differences in probability values, not simply the order of list, tuple, or dictionary items.  Running this harness using TSTL's {\tt
  --checkProcessDeterminism} flag required less than 10 seconds on
average to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm).  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue, 
required another 92 seconds.  Interestingly, the way that {\tt
  python-afl} is used in TSTL means that afl stability is 100\%,
because the {\tt PYTHONHASHSEED} has already been chosen before the
fork in afl executions.

>>>>

Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms.  The {\tt flatten\_and\_sort} function is needed because we care about actual differences in probability values, not simply the order of list, tuple, or dictionary items.  Running this harness using TSTL's {\tt
  --checkProcessDeterminism} flag required less than 10 seconds on
average to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm).  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue, 
required another 92 seconds.  Interestingly, the way that {\tt
  python-afl} is used in TSTL means that AFL stability is 100\%,
because the {\tt PYTHONHASHSEED} has already been chosen before the
fork in AFL executions.

### 2018-04-29 implementation.tex
AFL capitalization change

Because TSTL has a first-class interface to afl \cite{aflfuzz}, we can even use
afl's sophisticated heuristics to perform very thorough, week-long
checks for nondeterminism, using strategies built to find subtle
security vulnerabilities in C programs.  Equally importantly, since
there is substantial overhead in nondeterminism detection, afl can be
used to predict whether a program has potential nondeterminism; if the
afl \emph{stability} statistic (see the afl documentation for details
\cite{aflfuzz}) is lower than 100\%, it may indicate a problem.
Because of the idiosyncrasies of afl instrumentation and process
behavior, this is not always a reliable guide, but it is a very low
cost indicator produced as a by-product of fuzzing.

>>>>

Because TSTL has a first-class interface to AFL \cite{aflfuzz}, we can even use
AFL's sophisticated heuristics to perform very thorough, week-long
checks for nondeterminism, using strategies built to find subtle
security vulnerabilities in C programs.  Equally importantly, since
there is substantial overhead in nondeterminism detection, AFL can be
used to predict whether a program has potential nondeterminism; if the
AFL \emph{stability} statistic (see the AFL documentation for details
\cite{aflfuzz}) is lower than 100\%, it may indicate a problem.
Because of the idiosyncrasies of AFL instrumentation and process
behavior, this is not always a reliable guide, but it is a very low
cost indicator produced as a by-product of fuzzing.
